---
title: "SolCast Analysis of Blue Ridge Farm"
author: "Jordan White"
date: "9/16/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

We are interested in building a solar farm at Blue Ridge Rd in Chesapeake. This 
document is meant to analyze the physical parameters of such an installation. 
The data used here are solar irradiance and related data, all from a company 
called SolCast. Solcast uses satellite data combined with advanced meteorological 
models to estimate historical, current, and future irradiance of a given site. 
The error in their models tends to be low and is described in detail on 
their website: https://solcast.com/historical-and-tmy/validation-and-accuracy/ 
In brief, when compared to a ground measurement station that is close to the 
proposed site (a station in Hampton, VA), Solcast tends to overestimate 
irradiance by 1.14% and has an average error of less than 12% (Root mean squared 
error). The closest Solcast competitor has the same average bias and an average 
error that is over 2% greater.

Besides calculating solar irradiance, Solcast also calculates expected solar 
yields for utility scale installations. If a user uploads their site's power 
generation data, then Solcast's algorithms can learn how the site responds to 
cloud cover and other environmental factors. This produces better predictions 
for the given site and also allows for large maintenaince problems to be 
spotted. See this webpage for more information from Solcast: 
https://solcast.com/solar-radiation-data/inputs-and-algorithms/#pv-tuning-technology

```{r DataImport, echo=FALSE, warning=FALSE, message=FALSE}
library(dplyr)
library(readr)
library(lubridate)
library(zoo)
library(kableExtra)

setwd("~/Documents/SolarPower-BlueRidge/")
data <- read_csv("36.691364_-76.154219_Solcast_PT60M.csv")

# The below transform converts the UTC times to Eastern standard
data$PeriodEnd <- data$PeriodEnd - 4*60*60 # 4 secs scaled up to hours
data$PeriodStart <- data$PeriodStart - 4*60*60

# Also make a window averaged version, averaged by 6 hour windows
# Global Tilted Irradiation
convert <- zoo(data$GtiFixedTilt)
compressed_gti <- rollapply(convert, width = 6, by = 6, FUN = mean)
compressed_data <- tibble("GtiFixedTilt" = compressed_gti
                          ,"Time" = data$PeriodStart[seq(from = 3
                                                         ,to = length(data$PeriodEnd)-1
                                                         ,by = 6)])
convert <- zoo(data$GtiTracking)
compressed_gti <- rollapply(convert, width = 6, by = 6, FUN = mean)
compressed_data$GtiTracking <- compressed_gti
```

```{r Plotting_Irradiance, echo=FALSE}
# Plot tracking versus fixed tilt across the window averaged period
plot(compressed_data$Time
     ,compressed_data$GtiTracking
     ,type='l'
     ,ylim=c(0,1400)
     ,xlab="Time across four years (6 hour intervals)"
     ,ylab="Global Tilted Radiation (W/m^2)"
     ,main="Calculated Solar Irradiance\nof the Blue Ridge Farm"
     )
points(compressed_data$Time
       ,compressed_data$GtiFixedTilt
       ,type='l'
       ,col='#FF000095'
       )
legend('topright'
       ,c("Tracking Panels")
       ,pch = 22
       ,pt.bg = 'black'
       ,pt.cex=1.6
       ,bty='n'
       )
legend('topleft'
       ,c("Fixed Tilt Panels")
       ,pch = 22
       ,pt.bg = '#FF000095'
       ,pt.cex=1.6
       ,bty = 'n'
       )
```

The above figure is a calculation of what could have happened had we installed 
a solar system already. It calculates the power per square meter of solar panels 
across several years. The predictions are mostly based on solar irradiance and 
weather at the chosen location.

Looking at the above figure, you can see that the absolute amount of power we 
could generate depends on how many panels we could fit in the field. That in 
turn depends on their size and how big a shadow each row of panels would cast. 
Shadowing of one row of panels by another row is to be avoided as it 
can severely impact performance. 
If two or more panels are connected to the same DC/AC inverter, each panel will 
only output the power of the weakest panel in the link.

Let's consider a few possiblities just to gain some initial calculations. 
Given a panel that is 2 meters deep (300W Utility grade from Invensun), 
we could probably pack the 
rows of panels with 63.3 inches between them, assuming the average angle of the 
sun at this latitude. Assuming the worst case scenario, winter solstice, we 
could pack the panels with 154 inches between them. These are some wide 
numbers and the exact optimum for our location would require a search based 
algorithm that takes the above figure and a table of sun horizon and azimuth 
angles into account. I can write such an algorithm, given time, but let's start
with this set of best/worst case numbers.

We have perhaps 50 acres upon which we could build the panels. This is my low 
estimate that assumes we cannot use the full acreage, due to tree shading and 
installation space for inverters/batteries. The Inversun panels are 39.0551 
inches wide and will be about 63.3 inches deep if tilted to our latitude of 36.9ยบ. 
If each panel requires another 63.3 to 154 inches behind them, then the total 
area needed for each panel will be between 4,944.376 inches^2 and 8486.673 
inches^2. One square inch is $\frac{1}{1.5942*10^7}$ of an acre. Thus our min and max 
panel areas become $7.88*10^{-4}$ acres and $1.35*10^{-3}$. When divided across 
the 50 acres, we could install perhaps 37037 to 63451 300W panels.

Does that mean we'd get 300W times the number of panels? Not quite. 300W refers 
to the maximum output of the panels under ideal conditions. We have to 
consider the efficiency of the panels, which is likely around 20%. We also have 
to assume some loss of transmission efficiency, perhaps 10% loss. So our 
calculation will be $Number~~of~~Panels * m^2 ~~ per ~~ panel* 0.2 * 0.9 * Irradiance~~per~~hour$


```{r}
# Calculations using R code:
panel_area = 1.956*.992 # meters squared
low_end = 37037* panel_area*.2*.9
high_end = 63451*panel_area*.2*.9
fixed_tilt_low <- low_end*data$GtiFixedTilt # per hour basis
fixed_tilt_high <- high_end*data$GtiFixedTilt
tracking_low <- low_end*data$GtiTracking
tracking_high <- high_end*data$GtiTracking

#Summarize the MegaWatt/hours, remove the night hours (zero)
summary(tracking_high[tracking_high > 0])/1000000
summary(tracking_low[tracking_low > 0])/1000000
summary(fixed_tilt_high[fixed_tilt_high > 0])/1000000
summary(fixed_tilt_low[fixed_tilt_low > 0])/1000000
```

At 5 cents per kW/h our monetary returns could be plotted as below:

```{r ROI, echo=FALSE}
tracking_high_dolla <- tracking_high/1000*0.05
fixed_low_dolla <- fixed_tilt_low/1000*0.05

plot(data$PeriodStart,tracking_high_dolla
     ,type='l'
     ,lwd = 2
     ,xlab="Time by year (hourly)"
     ,ylab="Raw Return (Dollars)"
     ,ylim=c(0,1400)
     ,main="Predicted Returns by the Hour\nAcross Several Years")
points(data$PeriodStart, fixed_low_dolla
       ,type = 'l'
       ,lwd=2
       ,col="#FF000095")
legend('topleft', c("High Estimate w/Tracking")
       ,pch=22
       ,pt.cex=1.6
       ,pt.bg = 'black'
       ,bty='n'
       ,cex=.9)
legend('topright', c("Low Estimate w/Fixed")
       ,pch=22
       ,pt.cex=1.6
       ,pt.bg = '#FF000095'
       ,bty='n'
       ,cex=.9)
```

The below table sums up the predicted returns for each year. Note that I pulled 
these data in September of 2020; hence, without the last three months it has less 
predicted return. Focus on the '17 -- '19 results which all look relatively stable.

```{r ROI_by_year, echo=FALSE}
data$TrackingHighDollar <- tracking_high_dolla
data$FixedLowDollar <- fixed_low_dolla
data <- data %>% 
    mutate(Year = year(data$PeriodStart)) %>%
    group_by(Year)
totals <- data %>%
    summarize(High_Estimate = sum(TrackingHighDollar)
              ,Low_Estimate = sum(FixedLowDollar))
totals <- totals[totals$Year > 2016,]

kable(totals)
```
